---
documentclass: book
classoption: oneside
title: "OpenRI User Manual"
date: "`r Sys.Date()`"
vignette: >
 %\VignetteIndexEntry{OpenRI User Manual}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
output: 
  BiocStyle::html_document:
        toc: true
        toc_depth: 3
        number_sections: true
        theme: journal
        highlight: monochrome
        toc_float: true
        code_folding: hide
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="##")
knitr::opts_chunk$set(dpi=300)
knitr::opts_chunk$set(cache=FALSE)
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(fig_cap="yes")
knitr::opts_chunk$set(eval=F)
```


# Overview {#index}

```{r logo, fig.cap="The logo for the OpenRI resource.", fig.width=10, fig.height=3.8, echo=FALSE, eval=T, fig.align="center",out.width="40%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.logo.png")
grid.raster(img)
```

<hr>

## Unmet Needs

Most available interactions have focused on genes (or proteins), neglecting the study of regulatory interactions involving non-coding regions of the genome. As a result, there is a lack of dedicated resources and utilities in this area. The OpenRI resource addresses these needs by providing manual curation of gene-directed regulatory interactions across varied contexts, such as organs, tissues, and cell types. More importantly, its web-server supports data mining that encompasses the non-coding genome.


## Data Resource

The OpenRI resource is specially designed to be an open and scalable platform for investigating gene-directed multimodal regulatory interactions. Currently, these interactions are organised into three modalities: `Activity-By-Contact (ABC)` for enhancer-gene maps (see [this](https://www.ncbi.nlm.nih.gov/pubmed/33828297)), `Promoter Capture Hi-C (PCHiC)` (see [this](https://www.ncbi.nlm.nih.gov/pubmed/31501517)), and `Quantitative Trait Loci (QTL)` (see [this](https://www.ncbi.nlm.nih.gov/pubmed/34493866)).

Genes involved in regulatory interactions are annotated with a range of ontologies on `functions` ([Gene Ontology](https://www.ncbi.nlm.nih.gov/pubmed/33290552)), `phenotypes` ([Human Phenotype Ontology](https://www.ncbi.nlm.nih.gov/pubmed/33264411)), `diseases` ([Mondo Disease Ontology](https://www.ncbi.nlm.nih.gov/pubmed/31701156)), protein domains including `SCOP superfamily domains` ([SCOP](https://www.ncbi.nlm.nih.gov/pubmed/7723011)), `Pfam domains` ([Pfam](https://www.ncbi.nlm.nih.gov/pubmed/33125078)), and `InterPro domains` ([InterPro](https://www.ncbi.nlm.nih.gov/pubmed/33156333)). 


## Data Mining

In response to the growing availability of summary-level epi/genomic datasets from cohort-scale studies, OpenRI offers web-based data-mining capabilities that cover the non-coding genome. These capabilities include: 

- [Mining - Search] allows users to enter keywords to query the resource, which returns context-specific pages and gene-specific pages related to the query.

- [Mining - Browse] allows users to navigate through different contexts such as organs, tissues, and cell types to interactively explore their regulatory interactions.

- [Mining - RI2Genes] allows users to mine the resource to identify potential regulatory genes from their input data (see [Example Output](/app/examples/_tmp_RMD_RI2Genes.html)).

- [Mining - RI2Terms] allows users to mine the resource to identify potential regulatory genes and enriched ontology terms from their input data (see [Example Output](/app/examples/_tmp_RMD_RI2Terms.html)).


```{r, eval=FALSE, echo=FALSE}
cd ~/Sites/XGR/OpenRI-site/BOOKLET
################################################
R
library(tidyverse)

## bookdown::render_book + bookdown::gitbook
## use configuration in '_output.yml' and '_bookdown.yml'
bookdown::render_book("now.OpenRI_booklet.Rmd", c("bookdown::gitbook","bookdown::pdf_book","bookdown::tufte_html_book")[1])

q('no')
####################################################
```


# Web-server

## Frontpage

```{r app-front, fig.cap="The OpenRI web-server's landing frontpage features four key data-mining capabilities: `Search`, `Browse`, `RI2Genes`, and `RI2Terms`.", fig.width=10, fig.height=4.9, echo=FALSE, eval=TRUE, fig.align="center",out.width="100%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.app.front.png")
grid.raster(img)
```


## Mobile 

The OpenRI web-server has been designed to be mobile-first and responsive (`frontend`), while a next-generation Perl web framework `backend` is utilised to enable real-time browsing and mining of the OpenRI resource.

<hr>

```{r app-iphone, fig.cap="Screenshots of the OpenRI web-server visited via a smartphone. On the left is the frontpage, and on the right is the about page.", fig.width=10, fig.height=6.5, echo=FALSE, eval=TRUE, fig.align="center",out.width="90%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.app.iphone.png")
grid.raster(img)
```


# Mining - Search

The faceted search serves as a centralised mining hub within the OpenRI resource. It provides hyperlinks within the search results, supporting full-text queries for both contexts and genes. When searching for contexts, the search results are hyperlinked to context-specific pages; these pages provide the context information along with a tabular display for regulatory interactions between genomic regions and genes. Similarly, when searching for a particular gene, the search results provide hyperlinks to gene-specific pages; these pages contain comprehensive gene information, including its regulatory interactions (displayed separately for different modalities), and where applicable, annotated terms (of various ontologies).

<hr>

```{r search, fig.cap="The faceted search serves as a centralised data-mining hub within the OpenRI resource, offering interactive exploration.", fig.width=10, fig.height=5.4, echo=FALSE, eval=T, fig.align="center",out.width="100%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.search.png")
grid.raster(img)
```


# Mining - Browse

```{r browse, fig.cap="The by-context browse interfaces facilitate the retrieval of regulatory interactions by navigating through different contexts. This illustration showcases an exemplar using the context `C0001: A549_Roadmap (PMID33828297)`. The red hexagons with integers represent sequential clicks for navigation. `Top`: The browse page displays a list of contexts for regulatory interactions, including the `Activity-By-Contact (ABC)` modality. `Bottom-left`: The context-specific page for `C0001: A549_Roadmap (PMID33828297)` provides a comprehensive list of regulatory interactions between genomic regions and genes. Each gene is hyperlinked to its gene-specific page for further exploration. `Bottom-right`: The gene-specific page for the gene `ROCK1P1` presents its regulatory interactions separately for different modalities (`ABC`, `PCHiC`, and `QTL`). Each context is hyperlinked to its corresponding context-specific page.", fig.width=10, fig.height=8.8, echo=FALSE, eval=TRUE, fig.align="center",out.width="100%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.browse.png")
grid.raster(img)
```


# Mining - RI2Genes

## Input Interface

> **Input**

- `Step 1`: Enter a list of user-input genomic regions.

- `Step 2`: Specify the reference genome build. Ensure that the input genomic coordinates (`Step 1`) are consistent with the specified genome build (`Step 2`).

- `Step 3`: Select from a list of contexts available under different modalities (ABC, PCHiC, and QTL).

> **Output**

- [Example Output](/app/examples/_tmp_RMD_RI2Genes.html) includes interactive tables showcasing regulatory genes and their corresponding evidence. Additionally, a summary of the input data and the runtime (computed on the server side) is provided to the users.

<hr>


```{r RI2Genes-interface, fig.cap="The RI2Genes interface is displayed.", fig.width=10, fig.height=5.3, echo=FALSE, eval=TRUE, fig.align="center",out.width="90%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.RI2Genes.interface.png")
grid.raster(img)
```


## Interactive Results

- Under the `Input Information` tab, `Interactive table` lists user-input genomic regions.

```{r RI2Genes-results-input, fig.cap="A summary table of the user-input data displayed under the `Input Information` tab for exploration.", fig.width=10, fig.height=6.7, echo=FALSE, eval=TRUE, fig.align="center",out.width="95%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.RI2Genes.results_input.png")
grid.raster(img)
```

<br>
<hr>

- Under the `Output: Regulatory Genes` tab, `Regulatory Gene table` lists regulatory genes. Each gene is hyperlinked to its `gene-specific information page`. `Evidence table` provides information on how regulatory genes are identified, including the contexts and regulatory interactions associated with them.

```{r RI2Genes-results-genes, fig.cap="`Regulatory Gene table` within the `Output: Regulatory Genes` tab is depicted in this figure.", fig.width=10, fig.height=7.3, echo=FALSE, eval=TRUE, fig.align="center",out.width="95%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.RI2Genes.results_genes.png")
grid.raster(img)
```



# Mining - RI2Terms

## Input Interface

> **Input**

- `Step 1`: Enter a list of user-input genomic regions.

- `Step 2`: Specify the reference genome build. Ensure that the input genomic coordinates (`Step 1`) are consistent with the specified genome build (`Step 2`).

- `Step 3`: Select from a list of contexts available under different modalities (ABC, PCHiC, and QTL).

- `Step 4`: Specify the top regulatory genes and select from a list of ontologies to identify enriched ontology terms.

> **Output**

- [Example Output](/app/examples/_tmp_RMD_RI2Terms.html) includes interactive tables presenting regulatory genes (with evidence) and ontology terms, along with graphical plots for the enriched ontology terms. Additionally, a summary of the input data and the runtime (computed on the server side) is provided to users.

<hr>


```{r RI2Terms-interface, fig.cap="The RI2Terms interface is displayed.", fig.width=10, fig.height=6.2, echo=FALSE, eval=TRUE, fig.align="center",out.width="90%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.RI2Terms.interface.png")
grid.raster(img)
```


## Interactive Results

- Under the `Input Information` tab, `Interactive table` lists the user-input genomic regions.

- Under the `Output: Regulatory Genes` tab, `Regulatory Gene table` lists regulatory genes, each hyperlinked to its `gene-specific information page`. `Evidence table` provides details on how regulatory genes are defined, including the contexts and regulatory interactions involved.

- Under the `Output: Ontology Terms` tab, `Ontology Term table` presents the enriched ontology terms.

```{r RI2Terms-results-terms, fig.cap="`Ontology Term table` within the `Output: Ontology Terms` tab displays the list of the enriched ontology terms.", fig.width=10, fig.height=9.5, echo=FALSE, eval=TRUE, fig.align="center",out.width="95%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.RI2Terms.results_terms.png")
grid.raster(img)
```

<br>
<hr>

- Under the `Output: Plots` tab, two output plots, namely `Dotplot` and `Forest plot`, visually illustrate the enriched ontology terms. Additionally, downloadable PDF files are provided.

```{r RI2Terms-results-plots, fig.cap="The figure showcases the Dotplot presenting the enriched ontology terms.", fig.width=10, fig.height=9.3, echo=FALSE, eval=TRUE, fig.align="center",out.width="95%"}
library(png)
library(grid)
img <- readPNG("booklet.OpenRI.RI2Terms.results_plots.png")
grid.raster(img)
```
