---
title: ""
always_allow_html: yes
output:
  rmarkdown::html_document:
        toc: false
        toc_depth: 2
        number_sections: false
        theme: readable
        highlight: default
        toc_float: false
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results="asis")
knitr::opts_chunk$set(cache=FALSE)
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(fig_cap="yes")
```

```{r, eval=FALSE, echo=FALSE}
cd ~/Sites/XGR/OpenXGR-site
################################################
R
library(tidyverse)
## rmarkdown::render + bookdown::html_document2
rmarkdown::render("now.RMD_RI2Genes.Rmd", bookdown::html_document2(number_sections=F,theme=c("readable","united")[2], hightlight="default"), output_dir="/Users/hfang/Sites/XGR/OpenXGR-site/app/examples")

q('no')
################################################
```


# {.tabset .tabset-fade}


## **Input Information**
A runtime of ``r ls_rmd$runtime`` (``r format(ls_rmd$sT, "%a %b %d %X %Y")`` - ``r format(ls_rmd$eT, "%a %b %d %X %Y")``), time zone (``r Sys.timezone()``).

> **Interactive table** for user-input genomic regions (n=``r nrow(ls_rmd$data_input)`` analysed). The table includes three columns: 1st column `Input` for input genomic regions in the format of `chr:start-end` (the genome build `hg19`), 2nd `Chr` for the chromosome, 3rd `Start` for the start coordinate, 4th `End` for the end coordinate, and 5th `Length` for the region length (bp).

```{r}
data_input <- ls_rmd$data_input %>% mutate(chr=str_replace_all(dGR,':.*',''), start=str_replace_all(dGR,'.*:|-.*',''), end=str_replace_all(dGR,'.*-','')) %>% transmute(`Input`=dGR, Chromosome=chr, Start=start, End=end, Length=as.numeric(end)-as.numeric(start) + 1)

data_input %>% DT::datatable(options=list(pageLength=5,searchHighlight=T,dom='Bfrtip',buttons=c('csv','copy'), dom='t',scrollX=T), style='default', caption="", rownames=FALSE, escape=F, extensions=c('Buttons'))
```


## **Output: Regulatory Genes**
This tab features two output tables: `Regulatory Gene table` and `Evidence table`.

> **Regulatory Gene table** contains information on regulatory genes (n=``r ls_rmd$num_LG``, identified from input genomic regions), including the column `GScores` (ranged `from 1 to 10`) that quantifies the degree to which genes are likely regulated by input genomic regions. `Regulatory genes` are sorted by their gene scores and are hyperlinked to `gene-specific information page`. Please refer to **`Evidence table`** for details on context-specific evidence.

```{r}
filename <- ls_rmd$xlsx_LG

message(sprintf("'%s' '%s' (%s) ...", filename, getwd(), as.character(Sys.time())), appendLF=TRUE)

#tb <- openxlsx::read.xlsx(filename) %>% as_tibble() %>% mutate(`Regulatory genes`=sprintf("<a href='http://127.0.0.1:3080/OpenRI/gene/%s' target='%s'>%s</a>", Gene, Gene, Gene)) %>% transmute(`Regulatory genes`, GScores=signif(GScore,digits=4), Description)

tb <- openxlsx::read.xlsx(filename) %>% as_tibble() %>% mutate(`Regulatory genes`=sprintf("<a href='%sOpenRI/gene/%s' target='%s'>%s</a>", ls_rmd$host_port, Gene, Gene, Gene)) %>% transmute(`Regulatory genes`, GScores=signif(GScore,digits=4), Description)

tb %>% DT::datatable(options=list(pageLength=5,searchHighlight=T,dom='Bfrtip',buttons=c('csv','copy'), dom='t',scrollX=T), style='default', caption="", rownames=FALSE, escape=F, extensions=c('Buttons'))
```

<br>

> **Evidence table** displays information about the genomic regions (listed under the column `GR`) that overlap the input genomic regions (listed under the column `Input`). The overlapped regions are used to define regulatory genes (listed under the column `Regulatory genes`) based on the evidence provided (see the column `Evidence`). The column `Context` indicates the contexts: the prefix `ABC_` for contexts under the modality `ABC (Activity-By-Contact)`, the prefix `PCHiC_` for contexts under the modality `PCHIC (Promoter Capture Hi-C)`, and the prefix `QTL_` for contexts under the modality `QTL (Quantitative Trait Locus)`.

```{r}
filename <- ls_rmd$xlsx_LG_evidence

message(sprintf("'%s' '%s' (%s) ...", filename, getwd(), as.character(Sys.time())), appendLF=TRUE)

tb <- openxlsx::read.xlsx(filename) %>% as_tibble() %>% transmute(`Input`=dGR, GR, `Regulatory genes`=Gene, Context)

tb %>% DT::datatable(options=list(pageLength=5,searchHighlight=T,dom='Bfrtip',buttons=c('csv','copy'), dom='t',scrollX=T), style='default', caption="", rownames=FALSE, escape=F, extensions=c('Buttons'))
```
